Vigilantism and Institutions:\\ Understanding Attitudes toward Lynching in Brazil

### 2022-12-12 response-reviewers/response-reviewers.tex

\vspace{.5cm}

\noindent Dear Editor and Reviewers,

\vspace{.5cm}

Thank you for giving us the opportunity to revise our manuscript, ``Vigilantism
and Institutions: Understanding Attitudes toward Lynching in Brazil'' (Ms. No.
RAP-22-0164). We have made several changes to the article based on the feedback
from the editor and reviewers, and we believe the manuscript has improved
significantly as a result.

In particular, we have focused on addressing the scope and methodology of our
study, as highlighted by Reviewers 1 and 2. We have edited the Introduction and
added a new paragraph to the Conclusion to better explain the scope of our
work. We have also included more details about our methodology in section two
and in the Supplementary Materials. In response to Reviewer 2's suggestion, we
have removed the significance tests from the manuscript. We have also added a
footnote to the Supplementary Materials indicating that we randomized the order
of the experiments to reduce the risk of carryover effects.

Furthermore, we have added a discussion of how our findings compare with those
of other studies, as suggested by Reviewer 1. We have also included the results
of our subgroup analyses in the Supplementary Materials, again following
Reviewer 1's recommendation. Additionally, we have included all references
suggested by Reviewers 1 and 2 in the main body of the manuscript.

Overall, we believe that the manuscript is now stronger and more appealing to a
broader audience. We discuss the changes in more detail below. We are grateful
for the feedback and suggestions provided by the editor and reviewers, and we
hope that we have addressed all of their concerns. Thank you again for the
opportunity to revise our manuscript.

@@

\textbf{1)} The editor writes: ``I agree with R2 that the framing of the study
could be improved. You currently cite a number of interesting figures and basic
facts about lynching in Brazil in the start of the conclusion. I think this
would fit better in the introduction. R2 also argues that the reference to the
Blair et al study on greater costs of local violence is imprecise. I am
guessing you mean something along the lines of likely to affect a greater
number directly than terrorism or civil war, which I think is reasonable but
could be said more directly.''

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}

We have moved the facts about lynchings in Brazil to the Introduction and 
the section now provides more context for our experiments. The
second paragraph of the Introduction now reads as follows (page 02):
``\textit{One of the most serious forms of extrajudicial violence is lynching.
  Lynching can be defined as ``incidents of physical violence committed by
  large numbers of private citizens against one or more individuals accused of
  having committed a ``criminal'' offense, whether or not this violence
  resulted in the death of the victim(s)'' \citep[645]{godoy2004justice}.
  Although lynchings occur in more than one hundred countries in all regions of
  the world \citep{jung2020lynching,smith2019contradictions}, Latin America has
  been particularly affected by a sharp increase in vigilante violence.
  Lynching episodes have been reported in Guatemala, Mexico, Peru, Venezuela
  and other countries in the region \citep{barbara2015vigilantes,
  cruz2019determinants, godoy2004justice}. From 2011 to 2015, Brazil registered
  about 2,500 lynching episodes, and 173 people were killed by angry mobs in
  2015 alone---nearly one execution every two days
  \citep{barbara2015vigilantes, oliveira2016mob}. According to José de Souza
  Martins \citeyearpar{martins2015linchamentos}, who has studied lynchings in
  Brazil for more than thirty years, these figures are not only the highest in
  the country's history, but among the highest in the world. The people who
participate in lynchings are typically young men, but they also sometimes
include teenage women and girls, elderly women, and even members of the local
police \citep{moura2017linchamentos}.}

We have also changed the last sentence in the first paragraph to make our
claims more precise. It now reads (page 02): ``\textit{This
omission is significant not only because vigilantism can deepen group enmities
and lead to cycles of violence, but also because it undermines the legitimacy
of the state and the rule of law, which are essential for economic development
and democratic stability \citep{jung2020lynching, tankebe2009self}}.''

@@

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}

We have followed Reviewer 1's suggestion and have extended section two (``When
Is Lynching Perceived as More Justified?'') to provide further information
about our choice of covariates. The third paragraph of that section now reads
(page 04): ``\textit{In addition to our knowledge of the Brazilian case, we
  also chose our attributes based partly on existing work on crime,
  vigilantism, and extrajudicial violence. From the prison violence literature,
  we know that the relative age differences and the genders of perpetrators and
  victims affect how it is perceived \citep{fleisher2009myth}. We likewise know
  that certain offenses are seen as especially reprehensible
  \citep{skarbek2014social}, so we included such offenses, like molestation,
  alongside less serious offenses. We include race because, in the United
  States, violent lynchings have often been wielded for racist reasons
  \citep{dray2003hands}. Likewise, we include residency because extrajudicial
  violence is often used against people ``who don't belong'' or are
  ``outsiders'' in some respect. Finally, in honor cultures that value
  retaliatory violence, people believe that victims and the family of victims
  have a special right (and often an obligation) to enact retribution
\citep{weiner2013rule}.}

As we show in the next section, we have also included a new paragraph in the
Conclusion to discuss the relationship between our findings and the existing
literature on extralegal violence. By doing this, we aim to provide a more
comprehensive overview of our work and its significance within the broader
field.

@@

\noindent \textbf{4)} The editor writes: ``It would be useful to add a bit more
discussion of Brazil in comparative perspective. I wonder if the results for
Brazil are likely to be more general and that the US could be an outlier, for
example.''

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}

We have made the requested changes and significantly expanded the Conclusion
section based on the feedback from the editor and Reviewer 1. We are grateful
for this suggestion, as it has allowed us to better connect our article to the
broader literature on vigilantism.

The third paragraph of the Conclusion, now the longest in the section, is
exclusively dedicated to discussing the scope conditions of our research. We
have linked our findings to important works in the literature and have also
discussed how a culture of honor may help to explain some of our main findings.
This new paragraph provides valuable insight and enhances the broader
significance of our study. Overall, we believe that the revised Conclusion
section is stronger and more useful to readers interested in this topic.

@@

\vspace{.3cm}

\noindent \textbf{Response:}

\begin{quote}

Reviewer 1 raises an important issue. We agree that we had not adequately
explained the scope conditions of our research and have therefore added a new
paragraph to the Conclusion to provide additional context. In this paragraph,
we compare our results with studies on vigilantism in other Latin American
countries and the United States, and we also discuss the role of social norms,
particularly those related to a culture of honor, in shaping attitudes towards
lynching in Brazil. Additionally, we speculate on why race does not seem to
influence individual preferences for lynchings in Brazil.

@@

\noindent \textbf{2)} Reviewer 1 writes: ``With respect to experiment two---I
would be very clear about the control group (in my reading) having been
``treated'' by the conjoint. In terms of estimation this is fine, but in terms
of baseline attitudes that we might use to benchmark in future studies, I do
worry a bit.  I would also encourage a bit more description of the subgroup
analyses on experiment 2 since I think they are likely interesting/important
for future work.''

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}

We agree that our explanations regarding the experimental design were
inadequate. We actually randomized the order of the experiments to avoid
carryover effects, but we did not mention this explicitly in the text. We have
therefore added a footnote to the Supplementary Material to clarify this point.
The footnote, which can be found on page 83 of the SM, reads: ``\textit{To
prevent potential carryover effects caused by the conjoint, we randomized the
order of the conjoint and the information provision experiments
\citep{perreault1975controlling}.''} We believe that this procedure ensures
that our estimates are unbiased.

We agree that the subgroup analyses may be of interest to scholars. In response
to this feedback, we have added a short comment about each of our subgroup
tests to the Supplementary Materials. These comments provide information about
the statistical significance of the results and highlight any variations that
we observed according to the covariates that we controlled for. For example, in
Section C.4.1 of the Supplementary Materials, where we assess whether our
results for the conjoint experiment differ between genders, we write the
following (page 24): ``\textit{Results do not seem to vary according to the
  gender of the respondent. We focus here on the differences between males and
  females and exclude the 11 observations in which respondents preferred not to
say their gender or marked ``other'' in our questionnaire. Across all conjoint
experiment attributes, we see an overlap between the 95\% confidence intervals
for males and females.}'' Similarly, in Section C.4.3, in which we test whether
our results vary by race, we write (page 32): ``\textit{Below are our results
  when we disaggregate the data by race. We find that they are almost identical
  is all dimensions except for offense. Asian respondents are much less likely
to select profiles that contain pickpocketing as a crime.}''

All analyses are also accompanied by their corresponding \texttt{R} code to
facilitate the replication of our results.

@@

\noindent \textbf{2)} Reviewer 2 writes: ``The introduction motivates the study
by noting that ``local violence'' can be more costly than war or terrorism
violence. This raises questions about what is ``local violence'', and I'm not
sure the cited source after this statement clarifies the situation. The cited
Blair at al piece indicates that local violence is ``possibly'' more costly
than war or terrorism, but this seems imprecise to me. If there is more space
to defend this statement more, that might be ok, but it is probably better just
to cut it and motivate the study in other ways.''

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}

Reviewer 2's concern about the imprecision of our statement is valid. We have
deleted the original sentence and replaced it with the following (page 02):
``\textit{This omission is significant not only because vigilantism can
  entrench group animosities and create vicious cycles of violence, but also
because it severely undermines state legitimacy and the rule of law, two key
pillars of democratic stability and economic development
\citep{jung2020lynching, tankebe2009self}}.''

@@

\noindent \textbf{3)} Reviewer 2 writes: ``Why is the scale of disagree to
agree 0-100? Could the authors cite a source or sources on why this precise
range was used, and briefly discuss some tradeoffs with such a scale? Why not
a, for example, seven-point Likert scale?''

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}

Reviewer 2 is correct that we did not explain why we used a 0-100 scale in our
study. We chose this scale because it allows us to measure a wider range of
responses and capture the nuances of our treatments more accurately. This scale
is also easy to interpret and convert into discrete categories, which makes it
useful for our analysis. Additionally, the 0-100 scale is the default range
used by many survey platforms, such as
\href{https://www.surveymonkey.co.uk/curiosity/new-introducing-slider-star-rating-question-types/}{SurveyMonkey},
\href{https://www.questionpro.com/blog/use-slider-scales-for-a-more-accurate-rating/}{QuestionPro},
and
\href{https://www.qualtrics.com/support/survey-platform/survey-module/editing-questions/question-types-guide/standard-content/slider/}{Qualtrics},
which was the firm we used to conduct our survey experiments. As such, our
respondents are likely familiar with this scale.

We implemented the 0-100 scale following the recommendation of
\citet{carter2020}, who used the same range in their research design. While we
agree that a seven-point Likert scale would also be a good option, we believe
that the 0-100 scale is appropriate for our purposes.

@@

\noindent \textbf{4)} Reviewer 2 writes: ``Related to the above, the wording of
the instructions seem to suggest that respondents could select anywhere between
0 and 49 if they disagree; unless I am mistaken, the instructions do not
indicate, for example, that 0 indicates fully disagree and 100 indicates fully
agree. Were any more detailed instructions provided? Is there any evidence that
people fully understood the instructions in this regard?''

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}

Reviewer 2 raises a valid concern about the instruction for our second
experiment. While we acknowledge that we could have worded the instructions
better, we believe that respondents were able to understand them as intended.
As we mentioned earlier, Qualtrics uses the 0-100 scale by default in its
survey platform, so our respondents, who were recruited from Qualtrics' online
panel, are likely familiar with this scale. Additionally, we have plotted the
distribution of responses for the second experiment and have found that the
majority of respondents selected zero, which indicates that they do not agree
that lynchings are justified. This is consistent with the findings of our
quantitative text analysis, which estimates a latent Dirichlet allocation (LDA)
model and finds that one of the most common topics in the text responses is
``lynchings are not justified'' (page 81 of the Supplementary Materials).

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{hist.pdf}
  \caption{Support for Lynchings - Experiment 2}
  \label{fig:hist}
\end{figure}

We then cross-checked the distribution of responses with the qualitative
evidence we collected for the conjoint experiment, in which participants had to
justify their choices. When we filter our data to only include respondents who
selected zero on the 0-100 scale, we find the following justifications for
their choices (Portuguese version in parentheses):

\begin{itemize}

  \item ``\textit{It is not justified, but I am forced to express my opinion. We have a death here, but it is not justified}.'' (``não se justifica, mas sou obrigado a opinar. temos uma morte. mas não se jutifica'') 
  \item ``\textit{I only chose because I had to select a case, as stated. But I do not agree with either of the lynchings and I believe that justice should cover the gap that people use as an excuse for such practice}.'' (``escolhi somente pois deveria obrigatoriamente selecionar um caso, como enunciado. mas não concordo com nenhum dos dois linchamentos e acredito que a justiça deveria cobrir a lacuna que as pessoas utilizam como desculpa para tal prática'')
  \item ``\textit{I don't know... I selected one because I had to do so, but I am strongly against lynching. Eye for an eye, tooth for a tooth}.'' (``não sei bem... escolhi porque tinha que escolher, mas sou fortemente contra linchamento. olho por olho, dente por dente'')
  \item ``\textit{First of all, I want to make it clear that I do not condone either of the two cases. I think there should be no lynching, regardless of the crime. In this approach, I think case 2 is much more serious, because the victim is a child and the perpetrator is an adult, unlike in case 1}.'' (``à princípio quero deixar claro que não compactuo com nenhum dos dois casos. acho que não deve haver linchamento, independente do crime. nessa abordagem, acho que o caso 2 é bem mais grave, pois a vítima se trata de uma criança e autor de adulto, diferentemente do caso 1'')
  \item ``\textit{Either of the answers is immoral. Never, under any circumstances, is vigilante justice justified}.'' (``qualquer das respostas é imoral. nunca, em hipótese alguma a justiça pelas próprias mãos é justificada'')

\end{itemize}

There are several more such examples in the text. In contrast, when we include
only individuals who selected 100 on the scale, these are some of the
justifications mentioned in the first experiment:

\begin{itemize}
  \item ``\textit{A killer deserves the death penalty}.'' (``um assassino merece pena de morte.'')
  \item ``\textit{Because it is a murder: eye for an eye!}'' (``por tratar-se de um assassinato: olho por olho !'')
  \item ``\textit{There must be some punishment}.'' (``tem que haver punição.'')
  \item ``\textit{In both cases lynchings is appropriate}.'' (``nos dois casos cabe o linchamento.'')
  \item ``\textit{Because Brazil's Judiciary is s**t}.'' (``porque a justiça do brasil é uma merda'')
  \item ``\textit{Even if the criminal is indigenous, they must pay, you do not mess with a child}.'' (``mesmo que o criminoso seja indigena, tem que pagar, com criança não se mexe'')
\end{itemize}

All sentences above are included in the \texttt{q13\_text} variable in our
data. Although we cannot provide a statistical test to assess whether all
participants understood the instructions correctly, the evidence above suggests
that respondents did not find the scale confusing and that the quantitative
responses largely mirror our qualitative evidence. We appreciate Reviewer 2's
feedback and are grateful for the opportunity to address this concern.

@@

\vspace{.3cm}

\noindent \textbf{6)} Reviewer 2 writes: ``There is other recent survey work on
support for vigilantism that could be acknowledged, like Zizumbo-Colunga 2017.
On Brazil, there is Schuberth and others. Beyond survey work, there is a
growing line of work on vigilantes or autodefensas in other Latin American
countries, such as Peru or Mexico, that could be briefly discussed. For example
Kloppe-Santamaria on Puebla in JLAS or Trevizo on Mexico generally in LAPS.
This work seems relevant in many ways.''

\vspace{.3cm}

The reviewer is correct that there is a growing literature on vigilantism in
Latin American countries and the articles cited by the reviewer are indeed
relevant. We have added some of them to footnote number 01 (page 02), which now
reads: ``\textit{On vigilantism more generally, see
\citet{cohen2022collective}, \citet{schuberth2013challenging},
\citet{smith2019contradictions}, and \citet{zizumbo2017community}}.''

Paragraphs three and four of the Conclusion (pages 11--12), in turn, now read:
``\textit{Even in the American South, Whites and Blacks also lynched people of
  their own race \citep{beck1997race}. In this respect, our paper also
  highlights that lynchings in the Global South may be more strongly connected
with the idea of ``popular justice'' than with racial animus
\citep{kloppe2019lynching, martins2015linchamentos}.} [...] \textit{Finally,
lynchings have often been described as a response to low state capacity
\citep{trevizo2022mexico}, yet in some cases state agents actively incite or
engage in vigilante violence themselves \citep{arias2010violent}.}''

### 2022-12-12 appendix/appendix.Rmd
small changes to memo, article and appendix

We adhered to the ethical guidelines provided by the Institutional Review Board
at Brown University and APSA's Principles and Guidance. To facilitate
transparency, we comment here on a few aspects of our research design. First, we
worked with Brown University's IRB to reduce the risk of harm for participants
taking the survey. This included consultation with a cultural expert to inform
the phrasing of the survey. Second, respondents received compensation via
Qualtrics, which paid respondents directly after they completed the
questionnaire. Each subject in Qualtrics' online panel received the equivalent
of 2.5 USD in Brazilian Reals (local currency). Respondents who did not finish
the survey did not receive compensation. The compensation was appropriate for
the participant population. As of August 17, 2020, Brazil's monthly minimum wage
was BRL 1039, which amounted to 191 US dollars. Assuming 40 working hours per
week, the hourly minimum wage equals 1.19 US dollars. Our survey took about 20
minutes to complete and respondents received 2.5 USD, therefore subjects
received a monetary compensation that was 6 times higher than the local minimum
wage. Third, we do not see any potential or perceived conflicts of interest in
carrying out this research. We received a grant of USD 10,000 to conduct this
research from the Centre for the Study of Governance & Society at King’s College
London, which in turn receives support from the Templeton Foundation. We are not
aware of any conflicts of interest from either source. All of the code and data
used in this article have already been made publicly available.

>>>>

We adhered to the ethical guidelines provided by the Institutional Review Board
at Brown University and APSA's Principles and Guidance. To facilitate
transparency, we comment here on a few aspects of our research design. First, we
worked with Brown University's IRB to reduce the risk of harm for participants
taking the survey. This included consultation with a cultural expert to inform
the phrasing of the survey. Second, respondents received compensation via
Qualtrics, which paid respondents directly after they completed the
questionnaire. Each subject in Qualtrics' online panel received the equivalent
of 2.5 USD in Brazilian Reals (local currency). Respondents who did not finish
the survey did not receive compensation. The compensation was appropriate for
the participant population. As of August 17, 2020, Brazil's monthly minimum wage
was BRL 1039, which amounted to 191 US dollars. Assuming 40 working hours per
week, the hourly minimum wage equals 1.19 US dollars. Our survey took about 20
minutes to complete and respondents received 2.5 USD, therefore subjects
received a monetary compensation that was 6 times higher than the local minimum
wage. Third, we do not see any potential or perceived conflicts of interest in
carrying out this research. We received a grant of USD 10,000 to conduct this
research from the Centre for the Study of Governance & Society at King's College
London, which in turn receives support from the Templeton Foundation. We are not
aware of any conflicts of interest from either source. All of the code and data
used in this article have already been made publicly available.

### 2022-12-12 article/article.tex
small changes to memo, article and appendix

The regulation of violence is a crucial function of the state. When states
control violence effectively, markets, politics, and civil society can flourish
\citep{besley2011pillars, north2009violence}. However, states still struggle to
maintain the monopoly of force in much of the Global South. Scholars have
written extensively about how rebel organizations build support to challenge
government institutions, yet we know little about what motivates local
extrajudicial violence and vigilantism in the developing world
\citep{bateson2020politics}.\footnote{On vigilantism more generally, see
\citet{cohen2022collective}, \citet{schuberth2013challenging},
\citet{smith2019contradictions}, and \citet{zizumbo2017community}.} This
omission is significant not only because vigilantism can entrench group
animosities and create vicious cycles of violence, but also because it severely
undermines state legitimacy and the rule of law, two key pillars of democratic
stability and economic development \citep{jung2020lynching, tankebe2009self}.

>>>>

The regulation of violence is a crucial role of the state. When states control
violence effectively, markets, politics, and civil society can flourish
\citep{besley2011pillars, north2009violence}. However, many states in the
Global South still struggle to maintain a monopoly on the use of force.
Scholars have written extensively about how rebel organizations build support
to challenge government institutions, yet we know little about what motivates
local extrajudicial violence and vigilantism in the developing world
\citep{bateson2020politics}.\footnote{On vigilantism more generally, see
\citet{cohen2022collective}, \citet{schuberth2013challenging},
\citet{smith2019contradictions}, and \citet{zizumbo2017community}.} This
omission is significant not only because vigilantism can deepen group enmities
and lead to cycles of violence, but also because it undermines the legitimacy
of the state and the rule of law, which are essential for economic development
and democratic stability \citep{jung2020lynching, tankebe2009self}.
